### 2014-11-20 proposal.Rmd
Added parallel chains to stan model.

---
title: "Mortality Notes"
author: "J Camac"
date: "9 July 2014"
output: pdf_document
bibliography: data/refs.bib
csl: downloads/style.csl
---

###Background

Plant mortality and the processes affecting it are integral to understanding vegetation dynamics. However, few mechanistic models of vegetation dynamic explicitly simulate mortality [@Hawkes_2000], and those that have, do so using little process-based information. For example, mechanistic models have accounted for plant mortality by assuming mortality rates are constant through time and space (Thorton 2007m White 2000, Delbart et al 2010) or that they change when a threshold has been surpassed (e.g. when Sitch et al 2003, Sato et al 2007). These simplifications are a result of a lack of process-based information on what causes mortality rates to vary temporally, spatially and between species (Hawkes 2000, McDowell et al 2011). In part, this is because researchers have placed a greater emphasis on modelling plant growth (Hawkes 2000). But it is also due to the difficulties associated with modelling mortality rates.

Temporal patterns in plant mortality are highly variable and episodic (Burgmann 1994). Mortality rates also vary between species and are strongly influenced by multiple biotic and abiotic factors that may interact with one another (e.g. drought and herbivory: McDowell 2011). Because of this complexity, calibrating empirical models or validating mechanistic models of plant mortality requires datasets that monitor large numbers of species and individuals across their lifespan and across multiple biotic and abiotic gradients. More importantly, it requires the cause of death to be known for each individual. However, few, if any, data meet all these requirements. Rather, most available data comes from monitoring programs that are usually shorter than many plants lifespans (Bugmann 1996), and that focus on single species at single sites. Furthermore, because of the inherent difficulty of defining the cause of death, most monitoring programs do not delimit the cause of death for each individual.

Consequently, there are very few mechanistic mortality models. In a review of 61 woody plant mortality models, Hawke (2000) found that only 4 were mechanistic. In these models, mortality occurred when carbon supply was insufficient for growth (e.g. Friend et al. 1997, Bossel 1986, Weinstein et al 1991, Bugmann et al 1997). By contrast, the majority of mortality models (57) were empirical and included forest yield models (Guan and Gertner 1991), gap models (Dale et al 1985), plant-environment models (King and Grant 1996) and spatial models (Jeltsch et al 1997). These empirical mortality models spanned the range from deterministic (REFS) to stochastic (REFS) and typically included correlations of one or multiple, abiotic and biotic factors such as competition (REF) and climate (REF) as well as state variables such as age (REF), size (Hurst et al 2011, , V) and growth rate (REF).

More recently, plant traits have been incorporated into empirical models in order to examine inter-specific variability in mortality rates. These studies have suggest that a several traits may reduce mortality rates. For example, several studies have found that species with higher wood density exhibit reduced baseline mortality rates. This is thought to be due to higher wood density resulting in increased structural support, which in turn reduces the chance of breakage but also herbivory by wood boring herbivores. Other traits have been found to convey mortality advantages under particular hazards. For example, bark thickness has been correlated with greater survival after fire due to the bark insulating the sapwood from extreme temperatures. Or having a deciduous life form often allows for greater survival during prolonged drought by allowing plants to essentially hibernate during periods of high water stress, relative to evergreen species. However, the vast majority of these studies assume that the effects of traits remain do not interact with changing environmental conditions or individual state variables such as growth rate or size. This is despite theory suggesting otherwise. For example wood density is likely to matter more for an adult plant then a juvenile plant because adults are taller and are more prone to being knocked over by windthrow, whereas juveniles are likely to be more sheltered by surrounding taller vegetation.

###Aims
In this project I aim to address three knowledge gaps in relation to traits and mortality rates.

1) Examine whether the effects of wood density (and other traits) on mortality are growth & size dependent or independent.
2) Conduct a literature review summarising how plant traits influence mortality rates under common hazards (e.g. wind throw, drought, fire, disease, herbivory).
3) Build a process-based mortality model that allows trait effects to interact with the type of hazard as well as growth and size.

##1 - Growth dependent & independent effects of wood density

The effect of wood density on plant survival is often fitted to empiricial models under the assumption that it is independent of state variables such as growth rate and size (i.e. models with no interaction terms). Here, the theory is that higher wood density equates to lower mortality rates regardless of state variables such as growth rate and size. However, wood density may actually have both dependent and independent effects on plant mortality. For example, higher wood density may reduce mortality rates via greater stem structural support and decreased wood herbivory regardless of other state and climatic variables. However, it may also interact with growth rates with high wood density and low or negative growth rates traits such as wood density may actually density may have both the effect of wood density is unl

##Growth dependent & independent effects of wood density
# Current model may not be ideal. Partly because WD can never be zero and thus the model is unrealistic. Centering WD also won't work as it changes the form of the model by making it multiplicative.

Potential solution is to use a different form for wood density (e.g. csWD^c1)
Explore other model forms.
Sub-questions:
1) What is the best growth rate metric to use? dbh, basal area?, mass?
2) Can we partition the effects of size and growth?
3) What model best explains the effect mortality?
4) Think about size related models that encapsulate increases at small and large sizes.
RGR decreases with size. Incr.GR increases with size. By including both growth rate and size in the model we can determine whether size has additional affects that are not explained by growth.

###Example JAGS model with no random effects

```{r, eval=FALSE}
library('dplyr')
library('R2jags')
#Loads BCI data and attaches trait data
load('../data_BCI/output/bci.mainstem.Rdata')
bci.traits <- read.csv('../data_BCI/rawdata/traits/BCI_traits_20101220.csv')
names(bci.traits) <- tolower(names(bci.traits)) # lowers trait column names for merging
bci.traits$sp <- tolower(bci.traits$sp) # lowers species code names for merging
bci.mainstem <- merge(bci.mainstem,bci.traits[,c('sp','sg100c_avg')],by = 'sp') #only uses species trait data exists for.
rm(bci.traits) # trait db no longer needed

#Subsets data to only include species with traits and has a status record.
#Also removes couple of rows that have 10 yr census intervals.
#Then adds a sp.id column
bci.mainstem<- bci.mainstem %>%
  filter(!is.na(census.interval) & !is.na(sg100c_avg) & !is.na(dead.next.census) & census.interval< 6) %>%
  mutate(sp.id = as.numeric(factor(sp)))

  #Preps data for model by centering and scaling predictors and time
  data <- list(
  n_obs = nrow(bci.mainstem),
  #n_spp = length(unique(bci.mainstem$sp)),
  #pp = as.numeric(factor(bci.mainstem$sp), as.character(unique(bci.mainstem$sp))),
  wd = unique(bci.mainstem[c('sp', 'sg100c_avg')])[,'sg100c_avg'],
  #dbh = scale(bci.mainstem$dbh)[,1]/2,
  #dbh_gr = bci.mainstem$dbh.gr,
  #dbh_rgr = scale(bci.mainstem$dbh.rgr)[,1]/2,
  #basal_area = scale(bci.mainstem$basal.area)[,1]/2,
  basal_gr = scale(bci.mainstem$basal.area.gr)[,1]/2,
  #basal_rgr = scale(bci.mainstem$basal.area.rgr)[,1]/2,
  ln_census = log(bci.mainstem$census.interval), #logged for model
  died = bci.mainstem$dead.next.census)

#Actual model
model <- function(x) {
  for(it in 1:n_obs) {

    died[it] ~ dbern(p[it])
    cloglog(p[it]) <- max(-9999, min(9999,  ln_census[it] + log(cd0*exp(-cd1 * wd[spp[it]]) + cd2 * exp(-cd3 * wd[spp[it]] * dbh_gr[it]))))
    }

  #Priors
  cd0 ~ dlnorm(0, 0.0001) # Must be positive or large growth rates are undefined.
  cd1 ~ dnorm(0, 0.0001) # Can be positive or negative (but theory would suggest positive)
  cd2 ~ dlnorm(0, 0.0001) # Has to be positive or small growth rates not defined
  cd3 ~ dnorm(0, 0.0001) # Can be positive or negative

  }
mod<-     jags.parallel(model.file = model,
                        n.iter = 4000,
                        data= names(data), 
                        envir=list2env(data),
                        inits = NULL,
                        parameters.to.save = c('cd0','cd1','cd2','cd3'))
```

#Potential Stan Model
```{r, eval=FALSE}
library(parallel)
library(rstan)
data <- list(
  n_obs = nrow(bci.mainstem),
  #n_spp = length(unique(bci.mainstem$sp)),
  #pp = as.numeric(factor(bci.mainstem$sp), as.character(unique(bci.mainstem$sp))),
  wd = bci.mainstem$sg100c_avg,
  #dbh = scale(bci.mainstem$dbh)[,1]/2,
  #dbh_gr = bci.mainstem$dbh.gr,
  #dbh_rgr = bci.mainstem$dbh.rgr,
  #basal_area = scale(bci.mainstem$basal.area)[,1]/2,
  basal_gr = scale(bci.mainstem$basal.area.gr,center = FALSE)[,1]/2,
  #basal_rgr = scale(bci.mainstem$basal.area.rgr)[,1]/2,
  census_length = bci.mainstem$census.interval, #logged for model
  died = bci.mainstem$dead.next.census)
cat('
    data {
    int<lower=0> n_obs;
    int<lower=0, upper=1> died[n_obs];
    vector[n_obs] census_length;
    vector[n_obs] wd;
    vector[n_obs] basal_gr;

    }
    parameters {
    real<lower=0> c0;
    real c1;
    real<lower=0> c2;
    real c3;

    } 
    model {
    for (it in 1:n_obs)
    died[it] ~ bernoulli(inv_cloglog(log(census_length[it]) + log(c0 * exp(-c1 * wd[it]) + c2 * exp(-c3 * basal_gr[it]))));
    }
    ', file=(StanModel <- tempfile()))

system.time(BCImodelbasal.gr <- sflist2stanfit(mclapply(1:1,mc.cores=1,
		function(i) stan(file=StanModel, data=data, inter=500,seed=123, chains=1, chain_id=i, refresh=-1))))
```

2) Examine whether the effects of wood density varies with plant size.
3) Write a literature review summarising how plant traits influence mortality across multiple hazards (e.g. wind throw, drought, fire, disease, herbivory)
4) Build a process-based model based on what we have learnt from the above objectives.

### Hazard Hypotheses
1. The most common hazards that can result in plant death are all size dependent and include:
**Competition**: This hazard will most strongly affect smaller individuals because they are more at risk of being shaded compared to larger plants. Traits affecting the likelihood of competition resulting in death will be correlated with shade tolerance such as *LMA* and *wood density*.

**Herbivory**: This hazard will most strongly affect small individuals more susceptible to complete foliage loss or being consumed entirely. Traits affecting the liklihood of herbivory resulting in death will be correlated with herbivory preference (e.g. *leaf nitrogen*), resistance (e.g. *wood density*) and persistance (*resprouting capabilities*).

**Fire**: This hazard is likely to be size dependent with small individuals more suspectible to fire compared to larger plants. This is because for small/juvenile plants will are less likely to have developed fire-resistant traits. They are also lower in stature making their leaves exposed to both low severity and high severity fires. Traits affecting the likelihood of fire resulting in death will be correlated with fire-resistance (e.g. bark thickness) and persistance (e.g. resprouting)

**Drought/Extreme heat/ Extreme cold**: These hazard is likely to affect smaller plants moreso than larger plants.

###Factors influencing small individuals
```{r, echo=FALSE}
par(xaxs='i', yaxs='i', mfrow= c(3,2))
curve(1/(1+((1/0.99)-1)*exp(5*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by fire', lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.99)-1)*exp(7*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by drought',lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.99)-1)*exp(7*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by high temperatures',lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.99)-1)*exp(7*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by low temperatures',lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.99)-1)*exp(7*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by herbivory',lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.99)-1)*exp(15*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by light competition', lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))
```

###Factors influencing large individuals
```{r, echo=FALSE}
par(xaxs='i', yaxs='i',mfrow=c(1,3))
curve(1/(1+((1/0.02)-1)*exp(-10*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by wind fall', lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(0.9/(1+((0.9/0.2)-1)*exp(-7*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by falling neighbour',lwd=2)
axis(1, at=c(0,1.2), labels=c('Small', 'Large'))

curve(1/(1+((1/0.001)-1)*exp(-8*x)), ylim=c(0,1), xlim=c(0, 1.2), xaxt='n', ylab= 'Mortality rate', xlab='Time', main='Death by disease',lwd=2)
axis(1, at=c(0,1.2), labels=c('0', 'infinity'))
```

###Traits are likely to affect shape of these hazard rates
```{r, echo=FALSE}
par(xaxs='i', yaxs='i', mfrow=c(1,2))
curve(1/(1+((1/0.02)-1)*exp(-10*x)), ylim=c(0,1), xlim=c(0, 2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by wind fall')
curve(1/(1+((1/0.02)-1)*exp(-5*x)), ylim=c(0,1), xlim=c(0, 2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by wind fall', add=T, col='red', lty=2, lwd=2)
axis(1, at=c(0,2), labels=c('Small', 'Large'))
legend(0.7,0.2, legend = c('low wd', 'high wd'), col=c('black','red'), lty=c(1,2), bty = 'n', lwd=2, cex=2,)

curve(1/(1+((1/0.02)-1)*exp(-10*x)), ylim=c(0,1), xlim=c(0, 2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by wind fall')
curve(1/(1+((1/0.02)-1)*exp(-5*x)), ylim=c(0,1), xlim=c(0, 2), xaxt='n', ylab= 'Mortality rate', xlab='Size', main='Death by wind fall', add=T, col='red', lty=2, lwd=2)
axis(1, at=c(0,2), labels=c('Small', 'Large'))
legend(0.7,0.2, legend = c('low severity', 'high severity'), col=c('black','red'), lty=c(1,2), bty = 'n', lwd=2, cex=2,)
```

## Survival analysis

- Most people model survival curves
- However, here we are interested in determining how trait alter hazards.

The survival function is intimately related to the **hazard** or **instantaneous mortality function**,  denoted $\lambda(t)$, which gives the event rate at time $t$ conditional on survival until time $t$:

\begin{equation} \label{eq:lambda} \lambda(t) = \lim_{dt \rightarrow 0} \frac{\Pr(t \leq T < t+dt)}{dt\cdot S(t)} = -\frac{S'(t)}{S(t)}.\end{equation}

This means that:
\begin{equation}
Survival(t) = \lambda_1(t) \times \lambda_2(t) \times \lambda_3(t)... \times \lambda_n
\end{equation}

### Here I plan to model hazards by...
####EG Fire

\begin{equation}
\lambda(t) = f(age)^{-1} \times \int_{0}^{\infty} p(intensity|age) \times \Pr(\textrm{Death in fire of intensity}\, i) \, \textrm{d}i
\end{equation}

\begin{equation}
\Pr(Death)  = \alpha + \beta_1 \times Size + \beta_2 \times Resprouter + \beta_3 \times Bark thickness
\end{equation}

##OTHER STUFF
Below are some excerpts from William Morris's thesis on mortality of trees.

The **finite mortality rate**, $M$, ignoring recruitment, is the complement of the ratio of final populatio size, $N_T$, to initial population size, $N_0$, divided by the length of the observation period, $T$,

\begin{equation} \label{eq:FiniteMort}
M = 1 -\frac{N_T/N_0}{T}
\end{equation}

However, in order to infer an unbiased expected proportion of dying individuals for periods shorter or longer than that used to calculate finite mortality one must calculate the **instantaneous mortality**, $M_T$:

\begin{equation} \label{eq:InstMort}
M_T = 1-\textrm{exp}(-\lambda T)
\end{equation}

where

\begin{equation} \label{eq:lambda}
\lambda = - \frac{\textrm{log}_e(1-M)}{T}.
\end{equation}

This function is widely used to describe the loss of individuals from a population over time. Here, when $M_T$ and $\lambda$ are more or less equivilent mortality rates are low.

Calculating mortality rate is made more complicated when multiple censuses (a period of time in which mortality risk is being assessed, where census interval = $T$) are considered, partlicalrly when census interval lengths vary. In such cases equations \ref{eq:FiniteMort} and \ref{eq:InstMort} cannot be used because $T$ is not dixed and substituting $T$ with average interval length has been shown to underestimate mortality (Kubo et al., 2000). This is a particular problem with large-scale monitoring programs such as BCI where the period of time between revisits of individual trees varies within a signle census because there are too many individuals for observations to be made simultaneously. Datasets such as BCI can take more than a year to complete an entire census.

When $T$ varies within a mortality dataset, estimates of population mortality rate will be biased if this variation is not accounted for (e.g. by assuming a mean value for $T$ and using equation \ref{eq:InstMort}. In order to achieve an unbiased estimate of $M_T$ and $\lambda$ when census length varies, maximum likelihood (Kubo et al, 2000) or Bayesian (He, 2003) methods can be used, which allow $T$ to be considered a continous variable.

The methods so far assume $\lambda$ is fixed in space and time, between individuals and over the lifespan of an individual. Often these assumptions are not met because individual hazard rates vary. Heterogeneous hazard biases estimates of population mortality rates (Sheil & May 1996) and this bias is greater for longer census intervals (Zens & Peart 2003). The need for unbiased estimates of mortality, despite the hazard heterogeneity present in data from forests and woodlands, has lead to the development and application of methods that move beyond considering mortality at the population level and directly modelling hazard of individual trees (Zens & Peart, 2003).

The logistic regression family of methods are the most common individual-based approach used to analyse tree mortality data (see Breece, Kolb & Dickson, 2008, Carus 2010, Flewelling & Monserud, 2002). Logistic regression models consider tree death, $\textrm{Pr}(y_i = 1)$, as a series of binomially distributed data with the probability of observing a death as function of a set of covariates, $X_i$, and associated coefficients, $\beta$ related to data via the logit link function.

\begin{equation} \label{eq:logit}
\textrm{Pr}(y_{i} = 1) = \textrm{logit}^{-1}(\beta X_i)
\end{equation}

The logit function, $\textrm{log}_e(p/(1 - p))$, transforms a quantity bound within the unit interval, such as probability, $p$, to cary between negative infinity and infinity. By relating mortality to covariates that can explain differences in death rate, individual based methods such as logistic regression can reduce the bias found in population rate estimates when individuals have variable hazard (Zens & Peart, 2003).

The complementary log-log (cloglog) link, $\textrm{log}_e(-\textrm{log}_e(1-p))$, is another link that can be used to estimate mortality rates.
The advantage of the cloglog link is that it is equivalent to standard survival analysis... NEED MORE INFO. It also reduces bias due to long and variable census intervals (BY?).
By adding the census interval $T$ to equation \ref{eq:logit} and substituting the logit link with a cloglog link yields

\begin{equation} \label{eq:cloglog}
\textrm{Pr}(y_{i} = 1) = \textrm{cloglog}^{-1}(\beta X_{i} + \textrm{log}(T))
\end{equation}

#Random notes

### MULTI-LEVEL NOTES:

####From Hedeker et al 2014

Nice summary of why multi-level models are essential for accounting for non-independence in nested datasets.

"An important question is then to determine the degree to which covariates are related to substance use initiation. In these studies it is often of interest to model the student outcomes while controlling for the nesting of students in classrooms and/or schools. In analysis of such grouped-time initiation (or survival) data, use of grouped-time regression models that assume independence of observations [Thompson, 1977; Prentice & Gloeckler, 1978; Allison, 1982] is therefore problematic because of this clustering of students. More generally, this same issue arises for other types of clustered datasets in which subjects are observed nested within various types of clusters (e.g., hospitals, firms, clinics, counties), and thus cannot be assumed to be independent. To account for the data clustering, multilevel models (also called hierarchical linear or mixed models) provide a useful approach for simultaneously estimating the parameters of the regression model and the variance components that account for the data clustering [Goldstein, 1995; Raudenbush & Bryk, 2002]."

## TWO REVIEW PAPERS TO READ Pickles et al 1995 & Hougaard 1995 ###

Lee et al 1992 has developed continuous0time survival models. However, the application of these models to grouped or discrete-time survival data is generally not recommended because of the large number of ties that result.

Instead, models specifically developed for grouped or discrete-time survival data have been proposed.

Both Han & Hausman (1990) and Scheike & Jensen (1997) have described proportional hazards models incorporating a log-gamma distribution specification of heterogeneity. Furthermore, Ten Have (1996) developed a discrete-time proportional hazards survival model incorporating a log-gamma random effects distribution, additionally allowing for ordinal survival and failure categories. Ten Have & Uttal (1994) used Gibbs samplng to fit a continuation ratio logit model with multiple normally distributed random effects.

## IMPORTANT NOTE ON link function

Doksum & Gasko (1990) note, that large amounts of high quality data are often necessary for link function selections to be relevant. McCullagh (1980) notes that the link function choice should be based primarily on ease of interpretation.

## Multilevel model that treats each individual's survival time as a set of dichotomous observations indicating whether or not an individual failed in each time unti until a person either experiences the event or is censored. This type of model is extensively described in Singer & Willett (2003) & is illustrated in Reardon et al (2002).

This approach is particularly useful for handling time-dependent covariates and fitting non-proportional hazards models because the covariate values can change across each individual's Tij timepoints.

# REFERENCES