### 2016-07-04 ms/supp.Rnw
supp material

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Unifying intraspecific and interspecific variation in tropical tree mortality}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{James S Camac}
\author[a]{Mark Westoby} 
\author[b]{S Joseph Wright}
\author[a]{Daniel S Falster}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
\affil[b]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}

% Please give the surname of the lead author for the running footer
\leadauthor{Camac}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Add something significant here}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{J.C. designed and performed the analysis; D.F. conceived the idea, supervized the development of this work; S.J.W \& M.W. provided biological insight. All authors contributed to the writing of this manuscript.}

\authordeclaration{The authors declare that they have no competing financial interests.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Tropical trees $|$ wood density $|$ growth $|$ mortality $|$ interspecific}

\begin{document}
% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalized the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\subsection*{Modelling true growth}

We did this by first modelling the observed initial ($D_{i,1}$) and final diameter ($D_{i,2}$) of each individual, $i$, as random realisations from a normal distribution centerd on their respective true estimates (i.e. $\widehat{D_{i,1}}$ and $\widehat{D_{i,2}}$), with the standard deviation fixed at \Sexpr{obs_error_sigma} (i.e. the observed measurement error):

\begin{equation}\label{eq:true_growth_obs} 
\begin{aligned}
D_{i,1} &\sim \mathcal{N}(\widehat{D}_{i,1}, 0.75) \\
D_{i,2} &\sim \mathcal{N}(\widehat{D}_{i,2}, 0.75)
\end{aligned} 
\end{equation}

The initial diameter for each individual $i$, $\widehat{D}_{i,1}$, was sampled as a random realisation from a log-normal distribution:
\begin{equation} \label{eq:true_dbh1}
\widehat{D}_{i,1} \sim \log \mathcal{N}(\mu_{D_1}, \sigma_{D_1}),
\end{equation}
where $\mu_{D_1}$ is the expected mean of log initial diameter and $\sigma_{D_1}$ defines that variation around that mean. $\mu_{D_1}$ was estimated using a normal prior $\mathcal{N}(0, 2.5)$, which represents our prior expectation that the mean initial diameter would be greater than zero but less than 150 cm. $\sigma_{D_1}$ was estimated using a weakly informative half Cauchy prior centerd on zero and a scale parameter of 2.5.

The final diameter measurement for each individual $i$, $D_{i,2}$ was estimated as:
\begin{equation}\label{eq:true_dbh2}
\widehat{D}_{i,2} = \widehat{D}_{i,1} + \Delta D_{i,2} \, (t_2-t_1),
\end{equation}
where $\Delta D_{i,2}$ represents the annual true increase in diameter (cm) and $(t_2-t_1)$ represents the number of days between initial and final measurements divided by 365.25. We modelled $\Delta D_{i,2}$ as a random realisation from a log-normal distribution centerd on zero and a standard deviation of 1. This prior conveyed our expectation that annual diameter growth could not be negative and is unlikely to be greater than 7 cm/yr.

Using this model we then used \texttt{rstan}'s \textit{optimizing} function to maximize the joint posterior and extract point estimates of true initial diameter, $\widehat{D}_{i,1}$, true final diameter, $\widehat{D}_{i,2}$, and true annual diameter growth $\Delta D_{i,2}$. We then used annual diameter growth, $\Delta D_{i,2}$, and annual basal area growth: $\frac{(\pi \, 0.25 \,\widehat{D}_{i,2}) - (\pi \, 0.25 \, \widehat{D}_{i,1})}{(t_2-t_1)}$ in the subsequent mortality analysis.

\subsection*{Mortality model}
\subsection*{Model fitting}
Our observations are for survival outcomes $S_{i, t_2\rightarrow t_3}$ (0 = alive, 1 = died) for each individual $i$ between censuses at time $t_2$ and $t_3$. We model these observations as random realizations from a Bernoulli distribution
\begin{equation}
S_{i,t_2\rightarrow t_3} \sim \textrm{Bernoulli}(p_{i,t_2\rightarrow t_3}),
\end{equation}
where $p_{i,t_2\rightarrow t_3}$ is the probability individual $i$ dies in the interval $t_2$--$t_3$ *(eq. \ref{eq:mortality_model}). As we only have a point estimate of $\lambda_i(t)$ at $t=t_2$, the cumulative hazard from  $t_2$--$t_3$ is approximated as
\begin{equation}
  \int_{t_2}^{t_3} \lambda_i(t) {\rm d}t \approx \lambda_i(t_2) (t_3-t_2).
\end{equation}

As we have multiple hypothesized functional forms of $\lambda_i$ (e.g. equations \ref{base_haz}\textendash\ref{add_haz}), we describe here the fitting procedure of our most complicated model, of which all others may be considered special cases. The instantaneous rate of mortality is
\begin{equation}
\lambda_i = \left(\overbrace{\alpha_{s[i]} \, \exp\left(-\beta_{s[i]} \, X_i \right)}^{\text{Growth dependent}} +
\overbrace{\gamma_{s[i]}}^{\text{Growth independent}}\right) \times \overbrace{\delta_{t[i]}}^{\text{Census effect}}
\end{equation}  
In this model, $\alpha_s$, $\beta_s$ and $\gamma_s$ are allowed to vary by species as a function of both wood density (measured at species level) and a species random effects, e.g.
\begin{equation}\label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \left(\frac{\rho}{\rho_c}\right)^{\alpha_{1}},
\end{equation}
with similar formulations for $\beta_s$ and $\gamma_s$. Here $\alpha_{1}$ captures the effect of wood density $\rho$ on $\alpha_s$, while $\alpha_{0,s}$ captures any other species-level residual error not explained by wood density for species $s$ within which individual $i$ belongs. The form of sub-model in eq. \ref{eq:submodels} ensures that parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ remain positive. On a log scale this model equates to an additive linear model , e.g. $\log \alpha_s = \log \alpha_{0,s} + (\alpha_{1} - \log \rho_c) \, \log \rho$.

We standardized wood density via the transformation: $\frac{\rho}{\rho_c}$, where $\rho_c=0.6$ g cm${^-3}$. This operation is equivalent to centering a log-transformed $\rho$ around $\rho_c$ and enables mean hyper-parameters  to be interpreted as the value for a species with average wood density of 0.6 cm${^-3}$. Similarly, we also centered $X_i$ at the lower 5\% quantile for both diameter increment and area growth (0.172 and 0.338, respectively), meaning $\alpha_s$ should be interpreted as the hazard rate when growth rate was very low.

We modelled species random effects as random realisations from log-normal distributions:
\begin{equation}
\alpha_{0,s} \sim \log \mathcal{N}\left(\mu_{\log \alpha_0}, \sigma_{\log \alpha_0}\right),
\end{equation}
with similar formulations for $\beta_{0,s}$ and $\gamma_{0,s}$.

All three mean hyperparameters, $\mu_{\log \alpha_0}$, the average log species residual error at zero growth, $\mu_{\log \beta_0}$, the average species error on the exponential decay rate, and $\mu_{\log \gamma_0}$, the average baseline log species residual error, were estimated using normal distributions centered on zero with a standard deviation of 2.5.
\begin{equation}
\mu_{\log \alpha_{0}},\mu_{\log \beta_0},\mu_{\log \gamma_0} \sim \mathcal{N}\left(0, 2.5\right).\\
\end{equation}

Standard deviations associated with species random effects (i.e. $\sigma_{a0}$, $\sigma_{b0}$, $\sigma_{c0}$), were estimated using positive half Cauchy priors centered on zero with a scale of 2.5.
\begin{equation}
\sigma_{a0},\sigma_{b0},\sigma_{c0} \sim \textrm{Cauchy+}\left(0, 2.5\right).\\
\end{equation}

For models that did not include species random effects, single intercepts were estimated for each parameter using log normal distributions centered on zero with standard deviations of 2.5.

We estimated wood density effects (i.e. $\alpha_1$, $\beta_1$ and $\gamma_1$) using normal priors centered on zero and a standard deviation of 2.5:
\begin{equation} \label{eq:rho_effects}
\alpha_1, \beta_1, \gamma_1 \sim \mathcal{N}\left(0, 2.5\right).\\
\end{equation}

We accounted for possible variation in mortality rates across different census periods by allowing $\lambda_i$ to be scaled by $\delta_t$ for each census $t$. We estimated $\delta_t$ as a random realization from a log-normal distribution centered on zero and a variance parameter $\sigma_{\delta}$ defines the variation in census effects:

\begin{equation} \label{eq:census_effect}
\delta_t \sim \log \mathcal{N}\left(0, \sigma_{\delta}\right)\\
\sigma_{\delta} \sim \textrm{Cauchy+}\left(0, 2.5\right).
\end{equation}