### 2017-05-31 ms/ms.Rnw
Edits throughout
- refine and shorten text
- update PNAS style
- better formatting of Supp Matt
- update units in figures

\documentclass[9pt,twocolumn,twoside]{pnas-new}\usepackage[]{graphicx}\usepackage[]{color}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

% Enable cross referencing to Evidence document
\usepackage{xr}
\externaldocument{supp}

\title{Unifying intra- and inter-specific variation in tropical tree mortality}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{James S Camac}
\author[a]{Mark Westoby}
\author[b]{S Joseph Wright}
\author[a,c]{Daniel S Falster}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Department of Biological Sciences, Macquarie University NSW 2109, Australia}
\affil[b]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[c]{Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Science, University of New South Wales}

% Please give the surname of the lead author for the running footer
\leadauthor{Camac}

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{Tree mortality is a fundamental demographic process affecting forest dynamics and carbon cycling. Here, for the first time, we use observed survival records over a 15 year period from more than 180,000 individuals, to simultaneously estimate growth-dependent and growth-independent (i.e. stochastic) mortality across 203 tropical forest tree species. We found that growth-dependent mortality was the predominate factor influencing tree mortality rates at Barro Colorado Island. Furthermore, we found that wood density most strongly influenced mortality rates by decreasing growth-independent mortality. However, wood density only accounted for a small fraction of the overall species variability observed in growth-independent mortality rates, suggesting that there must be other species traits that strongly affect mortality. Which traits and how many of them, however, remains unknown.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{J.C. performed research; D.S.F. conceived the idea; S.J.W. collected data. All authors contributed to the writing of this manuscript.}

\authordeclaration{The authors declare that they have no competing financial interests.\\

Data deposition: Code and data reproducing the results in this paper are available at \href{https://github.com/traitecoevo/mortality_bci}{github.com/traitecoevo/mortality\_bci}.
}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Bayesian $|$ Growth rates $|$ Hazard rates $|$ Tropical trees $|$ Wood density}

% 250 words
\begin{abstract}
Tree death is a fundamental process driving population dynamics, nutrient cycling, and evolution within plant communities. While past research has identified factors influencing tree mortality across a variety of scales, these distinct drivers are yet to be integrated within a unified, predictive framework. In this study, we use a cross-validated Bayesian framework coupled with classic survival analysis techniques to derive instantaneous mortality functions for 203 tropical rainforest tree species at Barro Colorado Island (BCI) Panama. Specifically, we develop mortality functions that not only integrate individual, species and temporal effects, but also simultaneously estimate the relative contribution of growth-dependent and growth-independent hazards on the overall instantaneous mortality rate. We show that functions that partition mortality rates into growth-dependent and growth-independent hazards, use stem diameter growth rather than basal-area growth, and attribute the effect of wood density to growth-independent death, outperform alternative formulations. Moreover, we show that the effect of wood density -- a prominent trait known to influence tree mortality -- explains only a small proportion of the total variability observed between species, and that temporal variation in mortality was large. Lastly, our analysis indicate that growth-dependent processes are the predominant contributor to rates of mortality at BCI. Combined, this study provides a framework for predicting individual-level mortality in highly diverse tropical forests. More importantly, it highlights how little we know about the causes of species-level and temporal plot-scale effects needed to effectively predict tree mortality.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalized the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{R}ates of plant mortality are known to be highly variable among individuals, coexisting species, between forests and from year-to-year \cite{Kobe:1995tw,Wyckoff:2002ul,Poorter:2008iu}. This variation has considerable consequences for forest structure and dynamics. For example, death of a single large tree can transfer up to 20000 kg of carbon from live-to-decaying carbon pools\cite{Chave:2005hz}. Furthermore it creates a gap in the canopy that can restart a successional race, during which 10's-100's more plants may die while competing for a spot in the sun. In models of forest dynamics, variation in mortality rates have been found to have a larger impact on forest structure than variation in absolute growth rates\cite{Johnson-2016}. As such, improving our ability to understand the mortality process is a priority for making accurate predictions about forest and carbon dynamics as well as water or nutrient cycles; especially in an era of rapid environmental change.

Two difficulties arise when studying tree mortality in tropical rainforests. The first is that large population sizes and long periods of observation are required to make inferences into how various factors affect mortality rates\cite{Poorter:2008iu}. This requirement arises in part because the observable outcome -- alive vs. dead -- is one-step removed from the variable we ideally want to measure: $\lambda_i(t)$, the instantaneous rate of mortality for individual $i$ at time $t$, also called the ``hazard function''. In classic survival analysis\cite{Kleinbaum:2012tn}, the probability an individual dies between times $t_2$ and $t_3$ is a function of $\lambda_i(t)$:
\begin{equation} \label{eq:mortality_model}
p_{i,t_2\rightarrow t_3} = 1 - \exp\left(- \int_{t_2}^{t_3} \lambda_i(t) {\rm d}t\right).
\end{equation}
Here $\int_{t_2}^{t_3} \lambda_i(t) {\rm d}t$ is the ``cumulative hazard'' between $t_2$ and $t_3$\cite{Kleinbaum:2012tn}. The observed survival outcome $S_{i, t_2\rightarrow t_3}$ (0 = alive, 1 = died) is then a realisation of this probability
\begin{equation}
S_{i,t_2\rightarrow t_3} \sim \textrm{Bernoulli}(p_{i,t_2\rightarrow t_3}).
\end{equation}
Since trees are long-lived and we are trying to estimate the shape of a continuous function from binary data, large sample-sizes are required. As such, detailed studies of tree mortality have only recently become possible, with the accumulation of growth and survival data from repeat surveys spanning several decades in plots containing thousands of individuals\cite{Anderson-Teixeira-2015a}.

Another difficulty when studying mortality -- similar to that faced in other fields like medicine and engineering -- is determining the shape of hazard functions that skilfully predict patterns in mortality. Past research has identified a range of factors with significant effects on tree survival, including an individual's growth rate\cite{Kobe:1995tw,Wyckoff:2002ul}, traits\cite{Poorter:2008iu,Kraft:2010kq,Wright:2010fl,AubryKientz:2013dg}, and size\cite{Vieilledent:2010fv, Hurst:2011ky}. Generally, however, these influences have not been integrated into a common hazard function\cite{Hawkes:2000ib,Zens:2003jv,McDowell:2011dr}.

\section*{Towards a unified model of tree mortality}

A specific challenge in developing hazard functions for plants is to estimate the relative contribution of growth-dependent and growth-independent hazards on an individual's overall hazard. While plants die via many causes, these broadly fall into two categories. The first are growth-dependent hazards, where plants die because of insufficient carbon assimilation for growth and repair. The second are growth-independent hazards, where plants die because of stochastic events, irrespective of their growth rate, such as  windfall or fire. To capture this distinction, we assume an individual's hazard is the sum of growth-independent ($\lambda_{\textrm{b},i}(t)$) and growth-dependent ($\lambda_{\textrm{g},i}(t)$) rates:
\begin{equation} \label{eq:additive}
\lambda_i(t)  = \lambda_{\textrm{b},i}(t) + \lambda_{\textrm{g},i}(t).
\end{equation}

A further challenge in developing hazard functions for plants is to integrate the conflicting relationships observed within and across species between mortality and growth. Within species, mortality rates are lower for fast-growing individuals, presumably because those individuals have superior carbon budgets and are thus able to tolerate or repair diverse stresses\cite{Kobe:1995tw, Hawkes:2000ib, Keane:2001db}. Empirical studies broadly support this theory, with many indicating exponential declines in mortality with increased growth rate, $X_i(t)$ \cite{Kobe:1995tw, Wyckoff:2000un, Kunstler:2005bn, Vieilledent:2010fv}, such that with parameters $\alpha, \beta$:
\begin{equation} \label{eq:gd}
\lambda_{\textrm{g},i}(t) = \alpha \textrm{e}^{-\beta X_i(t)}.
\end{equation}

By contrast, across species, there is a strong trade-off between growth and survival, with individuals from faster-growing species  exhibiting higher mortality rates than individuals from slower-growing species\cite{Poorter:2008iu, Wright:2010fl}. This trade-off may arise via traits with antagonistic effects, such as wood density. Denser wood -- which is more expensive to build and thus slows growth rates -- conveys enhanced physical strength, reduced stem breakage\cite{vanGelder:2006exa,Chave:2009iy}, embolism\cite{Hacke:2001kj, Jacobsen:2005fx}, fungal or pathogen attack\cite{Augspurger:1984wx}, and ultimately reduced mortality \cite{Poorter:2008iu, Chave:2009iy, Kraft:2010kq, Wright:2010fl}.

\begin{figure}[!hb]
\centering
\includegraphics{figures/fig1.pdf}
\caption{Outline of methodology.
A) Our data consist of repeat measures of stem diameter ($D$) and status ($S$, alive or dead) for individual trees at specific census dates ($t_1$, $t_2$, $t_3$).
B) We consider three alternative hazard functions: 1) a baseline hazard, 2) a growth-dependent hazard; and 3) a function that combines both baseline and growth-dependent hazards. The parameters of the models are biologically interpretable: $\alpha$ defines the instantaneous mortality rate at low growth rate; $\beta$ reflects the sensitivity of mortality rate to changes in growth rate; and $\gamma$ is the asymptote, or baseline hazard. Combined $\alpha$ and $\beta$ capture growth-dependent mortality, while $\gamma$ captures growth-independent mortality (e.g. windfall, fire) that kill a plant irrespective of its growth rate. For each model form, we consider two alternative predictors of growth, $X$ (basal area and stem diameter growth), as well as allowing for species-level effects on the parameters $\alpha$, $\beta$ and $\gamma$.
C) Log-loss function highlighting that the further the predicted probability is from truth the higher the logloss.
D) The predictive skill of alternative models is evaluated via 10-fold cross validation. The entire dataset is split into 10 folds ($F_1$, ..., $F_{10}$). Alternative models were fit 10 times ($M_1$, ..., $M_{10}$), using different combinations of testing and training data. We then average the estimated log loss for each of these 10 model fits.}
\label{fig1}
\end{figure}

In this study, we attempt to reconcile intra-specific and inter-specific factors as well as partition instantaneous mortality rates into growth-dependent and growth-independent rates. We achieve this by developing a unified model that incorporates individual-level, species-level, and temporal effects, within a cross-validated Bayesian framework. To derive this unified model, we compare the skill of three alternative hazard functions ($\lambda_i(t)$; Fig. \ref{fig1}B) in predicting patterns of tropical tree mortality in 203 tropical rainforest trees at Barro Colorado Island (BCI), Panama. Beginning from eq.\ref{eq:additive}, the simplest model (Fig. \ref{fig1}B, left), assumes a single baseline, growth-independent hazard rate. The second model (Fig. \ref{fig1}B, middle) assumes the risk of dying eventually declines to zero as growth increases (eq. \ref{eq:gd}). The third model (eq. Fig. \ref{fig1}B, right) is the summation of the two previous models, which allows $\lambda_i(t)$ to decrease exponentially with increasing growth rate, but, unlike a standard negative exponential, asymptoting at some baseline hazard $\geq 0$.  Combined, these three models capture a variety of functional responses previously proposed, including effects represented in current vegetation models \cite{Moorcroft:2001ws,McDowell:2011dr}, which have not previously been systematically compared.

To derive the unified model, we asked the following questions:
1) Does mortality rates vary substantially between censuses, such that a specific census effect is needed?
2) Does a hazard function including both growth-dependent and independent terms outperform a function including only one of these?
3) Which growth measure (stem diameter or stem area increment) more skilfully predicts growth-dependent mortality?
4) Does including wood density (a species trait) improve model skill, and if so, does it affect parameters of growth-dependent or independent hazard or both?
5) What proportion of inter-specific variability in hazard is explained by wood density?
6) What proportion of total variability in hazard is explained by species, census, growth-dependent and growth-independent hazards?

Answering these questions involves comparing the  predictive skill of alternative models. To achieve this, we used 10-fold cross-validation (Fig. \ref{fig1}D), to systematically assesses each model's skill in predicting outcomes in novel data (i.e. not used in model calibration) \cite{Hooten:2015, Stone:1997tg}. Evaluating models in this way is computationally expensive, making it impossible to run all possible model formulations. We therefore fit models across five iterative stages of model development, corresponding to the six questions evaluated above, of gradually increased complexity (see Methods).

\section*{Results}

\subsection*{Mortality over time}
Comparing the 3 intervals between censuses from 1995 to 2010, we found tree mortality progressively decreasing over time. The highest proportion of trees death occurred between 1995\textendash 2000 (\Sexpr{prop_deaths(BCI_model_dataset_true_unique, 1)}\%) followed by 2000\textendash 2005 (\Sexpr{prop_deaths(BCI_model_dataset_true_unique, 2)}\%) and 2005\textendash 2010 (\Sexpr{prop_deaths(BCI_model_dataset_true_unique, 3)}\%). Consequently, when we allowed hazard rates to be scaled by census (i.e. adding term $\delta$) we observed a significant increase in predictive skill (i.e. Fig. \ref{fig2}A). Assuming the long-term average census effect ($\delta$) is 1, deaths were \Sexpr{census_effects(final_model,1)}\% higher in 1995\textendash 2000, and \Sexpr{census_effects(final_model,2)}\% and \Sexpr{census_effects(final_model,3)}\% lower in 2000\textendash 2005 and 2005\textendash 2010, respectively.

\begin{figure}[!h]
\includegraphics{figures/fig2.pdf}
\caption{Mean ($\pm$ 95\% credible intervals) logarithmic loss whereby lower values equal higher predictive skill.
A) Five sequential stages of model selection with each stage increasing in complexity. Null: null model ($\gamma$); Census: inclusion of census effects ($\delta_t$); Growth rate: the inclusion of growth rate. This includes two possible hazard functions: growth-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and baseline + growth-dependent hazard with census effects ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); WD: Inclusion of wood density effect ($\rho$) on $\gamma$; Species: the inclusion of species random effects $s$ on all parameters.
B) Change in logloss under all wood density\textendash parameter combinations. Colors represent growth measure used: white = no growth measure, grey = basal area growth, black = dbh growth. Symbols represent general functional form: null/baseline (circle), growth\textendash dependent hazard model (triangle) and model incorporating both baseline and growth\textendash dependent hazards (square). Note: errors are too small to see in panel A.}
\label{fig2}
\end{figure}

\subsection*{Growth hazard functions}
We found that a hazard function with both growth-dependent and growth-independent terms (i.e. $(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$), significantly outperformed both the baseline only (i.e.$\gamma \delta_t$) or growth-dependent only (i.e. $(\alpha e^{-\beta X_{i}})\delta_t$) functions (Fig. \ref{fig2}A). Moreover, we found that predictive skill was higher when using stem diameter growth over basal-area growth.

\begin{figure}[h!]
\includegraphics{figures/fig3.pdf}
\caption{Relationship between species baseline mortality rate and wood density. Points are estimated mean baseline mortality rates $\pm$ 95\% credible intervals for each of the 203 BCI species used in this study. Blue trendline with grey shading shows average ($\pm$ 95\% credible intervals) expected relationship.}
\label{fig3}
\end{figure}

\subsection*{Effect of wood density on mortality rates}
Including wood density ($\rho$) as an effect on either baseline or growth-dependent parameters significantly improved model skill relative to a hazard function without such effects (Fig. \ref{fig2} A,B). The most parsimonious model, with highest predictive skill, was that which attributed the wood density effect to the baseline hazard term ($\gamma$; Fig. \ref{fig2} B). Wood density was therefore negatively correlated with a species' baseline mortality rate (Fig. \ref{fig3}).

% This meant that fast growing individuals from a low wood density species had on higher average mortality rates (Fig. \ref{fig4} A,B), and thus higher annual probability of death (Fig. \ref{fig4} C,D), relative to fast growing individuals of high wood density species.

\subsection*{Proportion of variation explained}
By partitioning the modelled variation between our two random effects (i.e. species and census; Fig. \ref{fig5}A), we found that approximately \Sexpr{extract_prop_explained(param_variance_explained, I('species'))}\% of variation in predicted mortality rates was due to species differences, \Sexpr{extract_prop_explained(param_variance_explained, I('wood_density'))}\% of which was captured by wood density. By contrast, census effects accounted for approximately \Sexpr{extract_prop_explained(param_variance_explained, I('census'))}\% of the total predicted variation in mortality rates. When we partitioned modelled variation into the two independent hazards we found that the growth-dependent hazard (i.e. the part of the model allowed to vary by individual growth) accounted for \Sexpr{extract_prop_explained(param_variance_explained, I('growth_dependent'))}\% of the overall variation in mortality rates, and the remainder was attributed to the baseline hazard.

\subsection*{Mortality rates, maximum species size and species light requirements}
We found that both $\alpha$, the effect of low growth rate, and $\gamma$, the baseline hazard term, were weakly and positively correlated with gap index (our measure of species light requirement; Fig. \ref{fig6} A,C, E). By contrast, $\beta$ (the parameter that defines the exponential decay with increasing growth rate) was negatively correlated gap index. The combination of these correlations suggest that species which predominately recruit in gaps (i.e. higher gap index) were more prone to death due to both low growth and stochastic chance. They also required faster growth rates to achieve a given mortality rate relative to a species that recruits in non-gaps. We detected no strong correlations between estimated species parameters and their associated maximum DBH (Fig. \ref{fig6} B,D,F,H).

\begin{figure}[h!]
\includegraphics{figures/fig5.pdf}
\caption{Proportion of predicted variation explained.
A) Predicted variation explained by wood density (grey) as well as species and census random effects.
B) Predicted variation explained by growth\textendash dependent and baseline (stochastic) hazards.}
\label{fig5}
\end{figure}

\section*{Discussion}
Our Bayesian framework coupled with cross validation revealed that the most accurate and parsimonious model of tropical tree death was that which: 1) partitioned mortality into growth-dependent and growth-independent hazards; 2) used stem diameter growth rather than basal-area growth; 3) attributed the effect of wood density to growth-independent mortality; and 4) incorporated temporal varability. Moreover, we found that rates of tropical tree mortality varied substantially between species and that wood density, a species level functional trait, explained only a small proportion of the overall inter-specific variation.

The findings of this study provide empirical support for dynamic vegetation models that estimate mortality as the sum of growth-dependent and growth-independent hazards\cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. We show that regardless of growth measure, incorporating both hazards significantly improves model predictive skill. This is because the additive effects of both hazards explicitly allow models to account for two independent causes of plant mortality. The growth-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stressors (e.g. competition, parasitization, localized herbivory). By contrast, the growth-independent hazard (i.e. baseline hazard) accounts for deaths caused by stochastic events that arise irrespective of an individuals growth rate (e.g. windthrow).

Another advantage of partitioning mortality into growth-dependent and growth-independent hazards was that we could estimate the proportion of variation attributed to each. Like many other studies\cite{King:1994vg, Ruger:2011cv, Kunstler:2016km}, our analyses highlight the importance of light competition in influencing tropical tree demographic rates. Specifically, we found that the growth-dependent hazard accounted for \Sexpr{extract_prop_explained(param_variance_explained, I('growth_dependent'))}\% of the total predicted variability in mortality rates (Fig. \ref{fig5}). This indicates that carbon budget deficiencies were a major contributor to mortality rates at BCI between 1995 and 2010.

\begin{figure}[h]
\includegraphics{figures/fig4.pdf}
\caption{Predicted changes in instantaneous mortality rates (A,B), and annual mortality probability (C,D), with increasing growth rates. A,C) Estimated species curves for all 203 BCI species used in this study. B,D) The average ($\pm$ 95\% credible intervals) expected curve for a high (0.8) and low (0.3) wood density species.}
\label{fig4}
\end{figure}

Incorporating the effect of wood density on mortality rates also improved predictive performance. Our analyses revealed that the most parsimonious combination of wood density effects was when it was attributed to only the growth-independent hazard term. Specifically, high wood density species had lower baseline rates relative to low wood density counterparts. This finding corroborates the observed negative correlation observed between mortality and wood density reported elsewhere\cite{Kraft:2010kq,Chave:2009iy}. More importantly, our analyses support the theory that wood density predominately reduce mortality rates by decreasing a species vulnerability to growth-independent stochastic events such as windthrow, trampling and treefall\cite{vanGelder:2006exa,Chave:2009iy}.

While wood density effects are now being incorporated in mortality algorithms of many vegetation models\cite{Moorcroft:2001ws,Scheiter:2013ed}, our analysis indicate that such effects are likely to only capture a small proportion of the overall inter-specific variation (Fig. \ref{fig4}). This means that such models are likely to severely underestimate the true variability in mortality rates. Consequently, this underestimation is likely to manifest in biased estimates of carbon, water and nutrient dynamics, with flow on effects for ecosystem properties such as GPP and NPP\cite{Johnson-2016}.

\begin{figure}[h!]
\includegraphics{figures/fig6.pdf}
\caption{Posthoc correlations between estimated species parameters and measures of species light requirement (A,C,E,G) and species maximum DBH (B,D,F,H). Points are mean ($\pm$ 95\% credible intervals) for each of the 203 species used in this study. Blue line shows the average trendline as determined from standard linear regression. Gap index is a binary species level index of the light required for recruitment. Here low numbers indicate the species readily recruits under dense canopy and 1 indicates that a species only ever recruits in gaps.}
\label{fig6}
\end{figure}

The high unexplained inter-specific variation indicates that there are other species traits, besides wood density, that affect tree mortality rates. This unexplained variation may possibly be explained by a species' light demand and maximum height, which has recently been proposed as two major axes of inter-specific variation in tropical rainforests\cite{Poorter:2006tq}. While we did not model mortality as a function of these traits, our posthoc correlations (Fig. \ref{fig6}) revealed that some inter-specific variation was indeed correlated with light demand. Specifically, we found that light demanding species (i.e. those that typically recruit in gaps; high gap index) had higher baseline hazard rates and were more susceptible to dying as a result of low growth, relative to those that readily recruit in shade (i.e. low gap index), supporting findings by others\cite{Ruger:2011cv}. By contrast, we detected no correlations between a species maximum stem diameter and mortality rates, both contradicting\cite{Poorter:2006tq} and supporting\cite{Wright:2010fl} findings by others.

We recommend future research should focus on examining how other traits influence growth-dependent and growth-independent hazards and what combination of traits minimises unexplained species variability. Having said this, researchers should be mindful that the inclusion of additional traits comes at the cost of complexity, increased uncertainty (i.e. more parameters to estimate) and increased minimum knowledge; all of which diminish the interpretability of models, but also the applicability of predicting mortality rates in many data poor ecosystems. Because of this, we also recommend future research focus on developing more integrative correlates of plant mortality. For example, we may have captured more inter-specific variability in baseline hazards had we used a integrative measure of plant durability such as a trait that incorporates both wood density and stem architecture.

Furthermore, future research should focus on examining how climate variables influence both growth-dependent and growth-independent hazards, particularly as climate-driven mortality is increasing globally\cite{vanMantgem:2009bi, Mcdowell:2013wf}. Our analyses revealed a significant decline in mortality at BCI between 1995 and 2010. Unfortuantely, we were unable to statistically attribute this temporal change to particular climatic variables due to low temporal replication (N=3). However, the high numbers of deaths that occurred 1995 and 2000, were likely the product of a severe drought caused by an El-Nino in 1997\textendash 1998. Droughts are known to have significant impacts on tree mortality rates in these ecosystems\cite{Condit:1995wl, Mcdowell:2013wf}.

How the length between censuses affect the estimation of growth-dependent and growth-independent hazards should also be examined. Large census intervals are likely to lead to underestimation of growth-dependent mortality, and consequently, overestimate the relative contribution of growth-independent hazards. This is because if a census interval is too wide, it will fail to account for deaths attributed to rapid declines in growth caused by senescence (especially for short-lived plants) or events such as drought. Because our analysis estimated an individual's past growth from 5 year census intervals, and because it included short-lived species (i.e. those with high annual mortality rates; see Fig. \ref{fig4}) and a drought period between 1995 and 2000, our estimates of growth-dependent and growth-independent hazards are likely to be biased.

Lastly, research should also examine how different growth measures influence predictive skill of mortality models. We have no clear explanation as to why DBH growth outperformed basal area growth in terms of model predictive skill, and we are unsure how other common measures such as biomass growth\cite{Chave:2005hz} will compare.

Here we showcase a new framework for modelling tropical tree mortality that attempts to unifies empirical evidence from within and between species studies. This framework also provides an approach for partitioning mortality rates into growth-dependent and growth-independent hazards. Our findings reveal that while wood density is an important trait affecting mortality rates, we are still only capturing a fraction of the overall species variability in mortality rates.

\matmethods{
\vspace{-2em}
\subsection*{Data}
We derived plant mortality models using individual growth and survival data\cite{Condit:2012nz} collected from a relatively undisturbed 50-ha tropical rainforest plot on Barro Colorado Island (BCI), Panama (9.15\textdegree;N, 79.85\textdegree;W). The climate on the island is warm and rainfall is seasonal with falling between April and November. BCI is managed exclusively for field research by the Smithsonian Tropical Research Institute.

Within the 50-ha plot the diameter at breast height and survival status of all free-standing woody plants that were at least 1.3 m tall and had diameter $\geq$ 1 cm were recorded in 1981-1983, 1985, and every 5 years thereafter\cite{Condit:2012nz}. For the purpose of modelling mortality as a function of past growth, we discarded data collected in 1982 and 1985 because diameter measurements were rounded to the nearest 5 mm, whereas later censuses were rounded to the nearest millimetre. Consequently, we modelled tree mortality as a function of past growth for censuses 1995\textendash 2000, 2000\textendash 2005 and 2005\textendash 2010. We discarded species that do not exhibit secondary growth (e.g. palms and ferns), contained fewer than 10 individuals or did not contain an estimate of wood density. We also excluded individuals that: 1) did not survive at least two censuses (two being required to estimate growth rate); 2) were not consistently measured at 1.3 m above ground; 3) were multi-stemmed; 4) resprouted or \textit{``returned from the dead''}; or 5) were extreme outliers -- stems which grew more than 5 cm/yr or shrunk more than 25\% of their initial diameter. In order to maintain statical independence of measurements, we sub-sampled the dataset by randomly selecting one observation per individual out of a possible three censuses. In total, \Sexpr{n_inds(BCI_model_dataset_true_unique)} individuals of \Sexpr{n_spp(BCI_model_dataset_true_unique)} species were included in our analysis.

Wood density for each species was estimated by coring trees located within 15 km of the BCI plot, because increment borers are prohibited within the plot itself\cite{Wright:2010fl}. Cores were broken into pieces, each 5 cm long and specific gravity of each piece was determined by oven drying (100\textdegree C) and dividing by the fresh volume (as measured by water displacement). Wood density was then calculated as an area-weighted average (g cm$^{-3}$), where area refers to the annulus represented by each piece, assuming a circular trunk.

\subsection*{Estimating true growth}

Field measurements of diameter inevitably include observation error, which effect estimates of growth rates. In our dataset, \Sexpr{prop_neg_growth(BCI_model_dataset_true_unique)}\% of estimated growth rates were negative. To ensure our mortality model was not biased by these unlikely values we first applied a probabilistic model to estimate \textit{``true growth''}, taking into account measurement error and the distribution of growth rate across the community. Details of this model can be found in Supplementary Material S1.

\subsection*{Hazard functions}
We fit a series of models with increasing complexity. At each stage, the best model from the previous stage was taken as input for the next stage. In stage 1 we fit the simplest model (a constant) and compared to a same where hazard rate varied by census. Stage 2 involved comparing predictive performance of the three functions described in (Fig. \ref{fig1}B) using either stem diameter or stem area growth rate. This allowed us to determine: 1) whether predictive performance was substantially improved with increasing functional complexity; and 2) whether predictive performance varied by growth measure. In stage 3, we examined inter-specific variability by using the best model of stage 2 but allowing various combinations of parameters to vary as a function of species wood density. This allowed us to determine if wood density improved predictive capacity, and if so, whether it predominately improved predictive capacity via growth-dependent or growth-independent processes. Stage 4 also examined inter-specific variability, however, here the focus was on using the best model of stage 2 and estimating the total predictive gain obtained if all parameters were allowed to vary as a function of species ID. The last stage was fitting the best model, with species random effects and wood density to the entire dataset. This allowed us to examine the range of species hazard curves, how they are affected by wood density, census and posthoc covariates, and the proportion of predicted variation explained by species, wood density, census and growth-dependent and baseline hazard terms.

\subsection*{Model fitting}

For all models we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{StanDevelopmentTeam:2015uz}. We executed three independent chains and in all cases modelled parameters converged within 2000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. After discarding the first 2000 iterations as ‘burn in’, a further 2000 iterations were taken from the joint posterior.

\subsection*{Model evaluation}
Predictive skill was quantified by estimating the average log loss across 10-folds for held-out data, $\bar{\mathcal{L}}$ (Fig. \ref{fig1}). Logarithmic loss -- commonly known as log loss, $\mathcal{L}$, measures the skill of a model by penalizing incorrect predictions, based on how wrong the predicted probability is from the observed outcome, $S_i$ (Fig. \ref{fig1}C). Lower $\mathcal{L}$ implies greater skill. The average log loss across all individuals for the $k$th fold of held-out data, $\mathcal{L}_k$, is then
\begin{equation} \label{eq:logloss}
\mathcal{L}_k = - \frac1{N_k}\,\sum_{i=1}^{N_k}\big(S_i\,\log(p_{i,t_2\rightarrow t_3}) + ( 1 - S_i)\,\log(1 - p_{i,t_2\rightarrow t_3})\big),
\end{equation}
where $N_k$ refers to the number of observations in fold.

\subsection*{Posthoc correlations}

In order to calculate our gap index (our measure of a species light dependence) we also used annual canopy census data collected during 1985\textendash1990 and 1990\textendash1995. The canopy census recorded the presence of vegetation in six height intervals, 0–2, 2–5, 5–10, 10–20, 20–30 and 30 m every 5 m across the 50 ha. We calculated the minimum observed number of vegetation strata above 2 m all 5x5 cells in each census period. We then transformed this to a binary gap index by where 1 refers to no vegetation above 2 m and 0 referring to densely vegetated by subtracting the above values from 5 (i.e., 0 = dense, 5 = gap). To acknowledge the influence of gaps on surrounding cells' light availability, we then calculated a weighted focal sum using a 3 x 3 cell moving window, for which the central cell was assigned a weight of 8, and the eight neighbouring cells were each assigned weights of 1. The combined impact of vegetation density of neighbouring cells on light availability at the central cell was therefore assumed to be equivalent to the impact of vegetation density at the central cell.

}

\showmatmethods % Display the Materials and Methods section

\acknow{
J.C was supported by the Science and Industry Endowment Fund (SIEF; RP04-174). M.W was supported by a fellowship from the Australian Research Council (ARC). D.F. was supported by an ARC discovery grant. We thank Rich FitzJohn, Lachlan McCalman and Daniel Steinberg for computational advice; Anias Gibert etc etc.  Fig. 1a uses images by Tracey Saxby, IAN Image Library (http://ian.umces.edu/imagelibrary)}

\showacknow % Display the acknowledgements section

% \pnasbreak splits and balances the columns before the references.
% If you see unexpected formatting errors, try commenting out this line
% as it can run into problems with floats and footnotes on the final page.
%\pnasbreak

% Bibliography
\bibliography{refs.bib}
%\pnasbreak
\end{document}