### 2018-07-23 ms/manuscript.Rnw
some final changes to the document

\documentclass[9pt,twocolumn,twoside, lineno]{pnas-new}\usepackage[]{graphicx}\usepackage[]{color}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template

\title{Unifying intra- and inter-specific variation in tropical tree mortality}

\author[a,1]{James S Camac}
\author[b]{Richard Condit}
\author[a,2]{Richard G FitzJohn}
\author[c]{Lachlan McCalman}
\author[c]{Daniel Steinberg}
\author[a]{Mark Westoby}
\author[d]{S Joseph Wright}
\author[e]{Daniel S Falster}

\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, New South Wales, Australia}
\affil[b]{Field Museum of Natural History and Morton Arboretum, Chicago, Illinois, USA}
\affil[c]{Data61, CSIRO, Australia}
\affil[d]{Smithsonian Tropical Research Institute, Apartado 0843–03092, Balboa, Panama}
\affil[e]{Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Science, University of New South Wales, Sydney, New South Wales, Australia}

\leadauthor{Camac}

% Please add here a significance statement to explain the relevance of your work
% 120 words
\significancestatement{We build a mortality model that not only integrates intra and inter-specific variation but also, simulataneously estimates the two fundamental hazards faced by plants -- Death due to low carbon budget and death independent of carbon budget -- and how a species' light demand, maximum dbh and wood density affect each. We parameterise this model using over 400,000 observed survival records collected over a 15-year period from more than 180,000 individuals across 203 species. We show that carbon-dependent hazards are a significant contribtor to tree mortality at Barro Colorado Island. Moreover, we show that species traits -- wood density, maximum dbh and light demand -- all affect tree mortality rates but ultimately only account for a small fraction of the total variability among species.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{D.S.F. conceived idea; all authors designed research; J.S.C., D.S.F. \& R.G.F. performed analysis in discussion with L.M, D.S. \& M.W; R.C. \& S.W. collected data. J.S.C \& D.S.F. wrote the manuscript}

\authordeclaration{The authors declare that they have no competing financial interests.\\

Data deposition: Code reproducing the results in this paper is available at \href{https://github.com/traitecoevo/mortality_bci}{github.com/traitecoevo/mortality\_bci}. Data from BCI is available at DOIs \href{http://doi.org/10.5479/data.bci.20130603.}{10.5479/data.bci.20130603.}, \href{http://doi.org/10.5479/data.bci20140711}{10.5479/data.bci20140711}.
}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com. Present address: Centre of Excellence for Biosecurity Risk Analysis, School of Biological Sciences, The University of Melbourne, Melbourne, Victoria, Australia

\textsuperscript{2} Present address: Department of Infectious Disease Epidemiology, Imperial College London, Faculty of Medicine, Norfolk Place, London W2 1PG, United Kingdom
}

% Please include two to five keywords, separated by the pipe symbol
\keywords{Bayesian $|$ Growth rates $|$ Hazard rates $|$ Tropical trees $|$ Wood density}

% 250 words
\begin{abstract}
Tree death is a fundamental process driving population dynamics, nutrient cycling, and evolution within plant communities. While past research has identified a variety of factors influencing tree mortality at both individual and species scales, these distinct drivers are yet to be integrated within a unified predictive framework. In this study, we use a cross-validated Bayesian framework coupled with classic survival analysis techniques to derive hazard functions for 203 tropical rainforest tree species at Barro Colorado Island (BCI) Panama. Specifically, we develop an empricial mortality model that not only integrates individual, species, and temporal effects, but is also the first to simultaneously estimate the independent contributions of carbon-related hazards and carbon-independent hazards on the overall mortality rates. We show that functions that separate mortality rates into carbon-dependent and carbon-independent hazards, use stem diameter growth rather than basal-area growth, and incorporate a species' wood density, maximum dbh and light demand outperform alternative formulations. Moreover, we show that while these three traits do affect both carbon-dependent and carbon-independent hazards, they account for only a small fraction of the total variability observed among species. Combined, this study provides a statistical framework for predicting individual-level mortality in highly diverse tropical forests. It also highlights how little we know about the causes of species-level and temporal plot-scale effects needed to effectively predict tree mortality.
\end{abstract}

\dates{This manuscript was compiled on \today}
%\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\dates{This manuscript was compiled on \today}
%\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalized the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{R}ates of plant mortality vary widely among individuals within a species, among coexisting species, between forests, and from year-to-year \cite{Kobe:1995tw,Wyckoff:2002ul,Poorter:2008iu}. This variation has large consequences for forest structure and dynamics. The death of a large tree can transfer up to 20,000 kg of carbon from living to decaying carbon pools \cite{Chave:2005hz}, while also creating a gap in the canopy that restarts a successional race, during which 100's of plants will die while competing for a spot in the sun. Models of forest dynamics suggest variation in tree mortality rates can have larger impacts on forest structure relative than variation in growth rates \cite{Johnson-2016}. Improving our understanding of the mortality process is therefore a priority for accurately predicting changes in population, carbon and nutrient dynamics of forests; especially in an era of rapid environmental change \cite{Anderegg:2012ep,Mcdowell:2013wf}.

Two difficulties arise when studying tree mortality in tropical rainforests. The first is that large population sizes and long periods of observation are required to make accurate inferences into how various factors affect tree mortality rates \cite{Poorter:2008iu}. This requirement arises in part because the observed outcome -- alive vs. dead -- is one step removed from the variable we ideally want to measure: $\lambda_i(t)$, the mortality (or `hazard`) rate for individual $i$ at time $t$. In classic survival analysis \cite{Kleinbaum:2012tn, Zens:2003jv}, the probability an individual dies between times $t_1$ and $t_2$ is a function of both $\lambda_i(t)$, and the amount of time lapsed:
\begin{equation} \label{eq:mortality_model}
p_{i,t_1\rightarrow t_2} = 1 - \exp\left(- \int_{t_1}^{t_2} \lambda_i(t) {\rm d}t\right).
\end{equation}
Here $\int_{t_1}^{t_2} \lambda_i(t) {\rm d}t$ is the cumulative hazard an individual experiences from  $t_1$ and $t_2$\cite{Kleinbaum:2012tn}. The observed survival outcome $S_{i, t_1\rightarrow t_2}$ (0 = alive, 1 = died) is then a realisation of this cumulative hazard on a probability scale:
\begin{equation}
S_{i,t_1\rightarrow t_2} \sim \textrm{Bernoulli}(p_{i,t_1\rightarrow t_2}).
\end{equation}
Second, as trees are long-lived and we are estimating the shape of a continuous hazard function ($\lambda_i(t)$) from binary data, large sample sizes collected over long durations are required. Detailed studies of tree mortality have thus only recently become possible, with the accumulation of growth and survival data from repeat surveys spanning several decades in plots containing thousands of individuals \cite{Anderson-Teixeira-2015a}.

\section*{Towards a unified model of tree mortality}

Plants die because of many causes, however, these broadly fall into two main categories \cite{Hawkes:2000ib,McDowell:2011dr}. The first are carbon-dependent hazards, where plants die because of insufficient carbon assimilation for growth and repair. The second are carbon-independent hazards, where plants die because of chance events that are unrelated to their carbon budget, such as windfall or fire. An individual's total hazard at a given time point is therefore the sum of these two independent hazards. However, to date, most statistical models of tree mortality do not explicitly estimate the independent effects of these two fundamanetal hazards faced by plants. Instead, plant mortality is typically modelled using: 1) logistic regression (i.e. with logit link), which does not estimate a rate, but rather the log odds of death at a given time point; or 2) via a proportional hazards approach such as a Cox proportional-hazard model or complementary-log-log link binomial regression, both of which assume hazards are multiplicative, and therefore, not independent of one another \cite{Kleinbaum:2012tn}.

There are several distinct advantages of modelling tree mortality as a function of two additive, and thus independent hazards: 1) The model explicitly accounts for the independent effects of the two fundamental types of hazards faced by plants; 2) it allows for the examination of the relative contributions each hazard provides to the overall hazard rate; 3) it allows for examination of how various factors affect each hazard term, and because of this; 4) allows ecological theories of how factors such as species traits affect plant mortality to be empirically tested using commonly collected data.

A major challenge in developing hazard functions for plants is the need to integrate the conflicting relationships observed within and across species between mortality and growth. Within species, mortality rates are lower for fast-growing individuals, presumably because those individuals have superior carbon budgets and are thus able to tolerate or repair diverse stresses \cite{Kobe:1995tw, Hawkes:2000ib, Keane:2001db}. Empirical studies broadly support this theory, with many indicating exponential-like declines in mortality with increased growth rate, $X_i(t)$ \cite{Kobe:1995tw, Wyckoff:2000un, Vieilledent:2010fv}. By contrast, across species, there is a strong trade-off between growth and survival, with faster-growing species exhibiting higher mortality rates relative to slower-growing species \cite{Poorter:2008iu, Wright:2010fl}. This trade-off between growth and survival may arise via species traits with antagonistic effects, such as wood density. Denser wood -- which is expensive to build and thus slows individual growth --  reduces stem breakage \cite{vanGelder:2006exa,Chave:2009iy}, embolism \cite{Hacke:2001kj, Jacobsen:2005fx}, pathogen attack \cite{Augspurger:1984wx}, and thereby mortality \cite{Poorter:2008iu, Chave:2009iy, Kraft:2010kq, Wright:2010fl}. However, to date, these intra- and inter-specific effects have not yet been integrated into a single hazard function in such a way that it allows one to examine their effects on the two fundamental hazards faced by plants.

\begin{figure}[!hb]
\centering
\includegraphics{figures/fig1.pdf}
\caption{Outline of methodology.
A) Our data consist of repeat measures of stem diameter ($D$) and status ($S$, 0 = alive or 1 = dead) for individual trees at specific census dates ($t_1$, $t_2$, $t_3$).
B) We consider three alternative hazard functions: 1) a carbon-independent baseline hazard (i.e. average hazard), 2) a carbon-dependent hazard, where growth rate is a proxy for carbon budget; and 3) a function that combines both carbon-dependent and independent hazards. The parameters of the models are biologically interpretable: $\alpha$ defines the low growth rate (i.e. low carbon budget) effect on the overall hazard; $\beta$ reflects the sensitivity of mortality rate to changes in growth rate; and $\gamma$ is the asymptote (Panel B; right). In the most complex model (Panel B; right), $\alpha$ and $\beta$ capture carbon-dependent hazards, while $\gamma$ captures carbon-independent hazards that kill a plant, irrespective of its growth rate/carbon budget (e.g. windfall, fire). For each model form, we consider two alternative predictors of growth, $X$ (basal area and stem diameter growth), as well as allowing for species-level effects on the parameters $\alpha$, $\beta$ and $\gamma$.
C) Each model's skill in predicting observed outcomes ($S$) is quantified via the log-loss function (eq. \ref{eq:logloss}). Lower logloss means higher predictive accuracy.
D) The predictive skill of alternative models was evaluated via 10-fold cross validation. The entire dataset is split into 10 folds ($F_1$, ..., $F_{10}$). Alternative models were fit 10 times ($M_1$, ..., $M_{10}$), using different combinations of testing (1 fold; orange) and training (9 folds; green) data. Predictive capacity was assessed by averaging the log loss's obtained from the 10 test data predictions.}
\label{fig1}
\end{figure}

Here, we develop a tree mortality model that not only reconciles the conflicting growth-mortality relationships observed among individuals and between species, but also explicitly accounts for carbon-dependent and carbon-independent hazards. We achieve this by evaluating the following unified hazard function that incorporates individual-, species-, and temporal effects:
\begin{equation} \label{eq:model}
\lambda_{i}(t) = \bigg(\overbrace{\gamma_{s}}^{\text{Carbon independent}} + \overbrace{\alpha_{s} \, \textrm{e}^{-\beta_{s} \, X_{s,i}(t)}}^{\text{Carbon dependent}}\bigg) \times \overbrace{\delta_{t}}^{\text{Census}}.
\end{equation}
Eq. \ref{eq:model} allows for the additive, thus independent, effects of carbon-dependent and carbon-independent hazards, includes a negative exponential effect of individual's growth rate (our measure of carbon budget), and allows for mortality to vary among censuses, via the random effect $\delta_{t}$. Further, the parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ vary by species, $s$. Here we include the effects of three species traits thought to be major drivers of inter-specific variation in tropical tree mortality -- wood density ($\rho$), light rdemand ($\upsilon$) and maximum dbh ($\psi$). We also include species random effects in order to capture species-level differences that are not otherwise accounted for by species traits (for details see Methods).

\begin{figure*}[!ht]
\includegraphics{figures/fig2.pdf}
\caption{Predictive skill of alternative hazard functions. Values are mean ($\pm$ 95\% confidence intervals) logarithmic loss, with lower values implying greater predictive skill and a logloss $\approx$ 0.693 indicating a model unable to distinguish an observed death from a survival.
A) Six sequential stages of model selection with each stage increasing in complexity. \textbf{Stage 1:} Baseline: constant ($\gamma$) vs Baseline with census effect ($\gamma\delta_t$); \textbf{Stage 2:} Best model of stage 1 compared to hazard functions that also account for individual growth rate (i.e. carbon budget). This includes two possible hazard functions: a carbon-dependent only hazard ($(\alpha e^{-\beta X_{i}})\delta_t$) and a baseline + carbon-dependent hazard with census effects ($(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$); \textbf{Stage 3:} The effect of a species' maximum dbh ($\psi$), wood density ($\rho$) and light demand ($\upsilon$) on $\alpha$, $\beta$ and $\gamma$ of most predictive model from stage 2; \textbf{Stage 4:} 2-trait combinations; \textbf{Stage 5:} three-trait combination; \textbf{Stage 6:} Maximum predictive performance by accounting for all species differences via the inclusion of species random effects, $s$ on all three parameters. Symbols represent the three hazard functional form: baseline (circle), carbon-dependent (triangle), and hazard containing both carbon-independent and carbon-dependent hazards (square).}
\label{fig2}
\end{figure*}

To validate this model, we also fit two alternative hazard functions: 1) a baseline hazard (i.e. a null model) and 2) a carbon-dependent only hazard. We then compare each functions ability to predict the survival status of \Sexpr{n_inds(BCI_training_full)} individual trees from \Sexpr{n_spp(BCI_training_full)} species recorded over three 5-year census periods at Barro Colorado Island (BCI), Panama (Fig. \ref{fig1}). The data are repeat censuses of stem diameter and tree status (alive vs. dead) taken approximately at 5-year intervals over a 15 yr period (Fig. \ref{fig1}A). In total \Sexpr{n_obs(BCI_training_full)} observations were used to fit these models. We compare the skill of different hazard functions in predicting outcomes in novel data (i.e. not used in model fitting) via 10-fold cross-validation (Fig. \ref{fig1}D) \cite{Hooten:2015, Stone:1997tg}. Evaluating models in this way is computationally expensive, making it impossible to run all possible model formulations. We therefore fit models across six iterative stages of model development.

In stage 1, we first ask whether average mortality rates vary substantially among censuses, and thus, whether a census effect is required. We achieved this by fitting a constant, $\gamma$, that is invariant with respect to species, growth rate, or census period (i.e. a null model) and compared it to a constant that is scaled by a census random effect, $\gamma\delta_{t}$. This stage revealed that there was significant variation in mortality rates observed among censuses, and as such a random effect for census was included as a scaling factor for all subsequent models.

In stage 2, we assessed the predictive performance of three alternative hazard functions (Fig. \ref{fig1}B). The simplest form (Fig. \ref{fig1}B, left) assumes a constant hazard. The second form (Fig. \ref{fig1}B, middle) assumes the risk of dying declines towards an asymptote of zero as growth, and thus carbon budget, increases. The third form (Fig. \ref{fig1}B, right) is the summation of the two previous models; this allows $\lambda_i(t)$ to decrease exponentially with increasing growth rate, but, unlike the second functional form examined here, asymptotes at some carbon-independent baseline hazard $\textgreater 0$. Combined, these three models capture a variety of functional responses that are both used in empirical studies of plant mortality \cite{Kobe:1995tw,Wyckoff:2000un} and are represented in current vegetation models \cite{Moorcroft:2001ws,McDowell:2011dr}, but which have not previously been systematically compared. In doing this, we also investigate which growth measure (stem diameter, stem area increment) more skilfully predicts carbon-dependent mortality.

In the remaining stages, we take the most predictive hazard function and growth rate measure from stage 2 and examine how species traits: wood density, light demand and maximum dbh, affected model predictive performance. Specifically, we examined each trait individually on all three $\alpha$, $\beta$ and $\gamma$ (stage 3), as well as 2-trait (stage 4) and 3-trait (stage 5) combinations on these three parameters. Lastly, we fit a model that included a species random effect on each parameter (stage 6). This allowed us to 1) examine the maximum predictive capacity when accounting for species differences; and 2) provides a relative indication on how much of the total species variation is captured by these three traits. Because these models were computationally expensive to run (with some models requiring over 400 computing hours), we were unable to examine interactions among traits or determine the most predictive combination of traits effects and parameters in a cross validated framework.

\section*{Results}

\subsection*{Mortality over time}
Examining the three 5-year intervals between censuses from 1995 to 2010, we found average mortality rates declining over time. The highest proportion of tree death occurred between 1995\textendash 2000 (\Sexpr{prop_deaths(BCI_training_full, 1)}\%) followed by 2000\textendash 2005 (\Sexpr{prop_deaths(BCI_training_full, 2)}\%) and 2005\textendash 2010 (\Sexpr{prop_deaths(BCI_training_full, 3)}\%). Consequently, when we allowed hazard rates to be scaled by census (i.e. adding the term $\delta_t$) we observed a small increase in predictive skill (Fig. \ref{fig2}A). Table S1 gives individual census effects.

\subsection*{Hazard functions}
Comparing the three hazard functions summarised in Fig. \ref{fig1}B, we found that the third function --- the function containing both carbon-dependent and carbon-independent hazard terms, i.e. $(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$ --- significantly outperformed both the baseline and carbon-dependent only hazard functions (Fig. \ref{fig2}). Moreover, we found that predictive skill was higher when using stem diameter growth relative to stem-area growth (Fig. \ref{fig2}A). Table S1 gives a summary of mean hyper parameter estimates.

\begin{figure*}[ht!]
\includegraphics{figures/fig3.pdf}
\caption{Estimated relationships between between species traits (maximum dbh, light demand and wood density) and model parameters $\alpha$ (low growth effect), $\beta$ (growth-mortality decay rate) and $\gamma$ (baseline/carbon-independent hazard). Points are the estimated mean ($\pm$ 95\% credible intervals) parameters for each of the 203 Barro Colorado Island species used in this study. Blue trendline with grey shading shows average ($\pm$ 95\% credible intervals) expected relationship across observed trait ranges.}
\label{fig3}
\end{figure*}

\subsection*{Species traits}
Using the hazard function form with highest predictive skill (i.e. $(\alpha e^{-\beta X_{i}} + \gamma)\delta_t$; Fig.~\ref{fig1}b right), we examined how effects of a species' maximum dbh ($\psi$), light demand ($\upsilon$) and wood density ($\rho$) placed on $\alpha$, $\beta$ and $\gamma$ affected model predictive accuracy. In general we found that including trait effects on these parameters decreased logarithmic loss, and thus, increased predictive accuracy (Fig.~\ref{fig2}). Moreover, we found that incorporating the effect of all three traits on these parameters resulted in higher predictive capacity relative to models that only incorporated one or two of the traits. However, this increase in predictive accuracy was small relative to including species random effects on each parameter, which account for unexplained species differences, signifying that substantial inter-specific variation in mortality rates remains unexplained (Fig. \ref{fig2} Stage 6). See Figs. S3-S5 for species-level parameter estimates.

\begin{figure}[h!]
\includegraphics{figures/fig4.pdf}
\caption{Predicted changes in annual mortality probability with increasing growth rate. A) Estimated average curves for all 203 Barro Colorado Island species used in this study. B,C,D) represent average ($\pm$ 95\% credible intervals) curve for a high (red) and low (blue) trait values. B) wood density values 0.2 vs 0.8 (g cm$^{-3}$); C) maximum dbh values 3 vs 180 cm; D) light demand values 0.5 vs 0.8. Note that the intercept is the sum of $\alpha$ and $\gamma$ (i.e. the stochastic chance of death plus effect of low growth.}
\label{fig4}
\end{figure}

In order to examine how species traits -- maximum dbh, light demand and wood density -- affect tree mortality rates, we fitted a final model containing all three species traits and a species random effect (which captures inter-specific variation that unexplained by traits) on each of $\alpha$, $\beta$ and $\gamma$. We found that each trait influenced tree mortality in different ways (Fig. \ref{fig3}-\ref{fig4}). Specifically, we found that light demanding species were more susceptible to low growth rates (i.e. higher $\alpha$), had higher growth-independent hazards (i.e. higher $\gamma$) and exhibited shallower declines in growth-dependent hazards with increasing growth rates (i.e. lower $\beta$). By contrast, we found that wood density had significant impacts on $\alpha$ and $\gamma$, such that species with high wood density were less susceptible to low growth rates and had lower carbon-independent hazards. Species' maximum dbh only had significant effects on $\gamma$, whereby species that can achieve larger dbh exhibited lower carbon-independent hazards. Table S1 for trait effects on each parameter.
\subsection*{Parameter relative importance}

Using the final model, we then examine: 1) its ability distinguish between individuals that died and individuals that survived via AUROC; and 2) the amount of inter-specific variation the model explains (i.e. $r^2$) relative to the proportions of individual deaths observed per species between 1995 and 2010 (Table.~\ref{table:1}). We found that the full model had an AUCROC of \Sexpr{table1[[1,3]]}, indicating that the model had a fair ability to correctly discriminate observed deaths from survivals. Moreover, we found that the full model also explained \Sexpr{table1[[1,2]]*100}\% of the total variation observed among species.

<<results="asis", echo=FALSE>>=
Hmisc::latex(table1,file = "", booktabs = TRUE, rowname = NULL,  
               colnamesTexCmd="bfseries",
               col.just = c(rep("c", 3)), label = "table:1",
               where = "!h",
               caption ="Relative effects of parameter removal on both $r^2$ and AUROC. $r^2$ estimates are derived 
               from modelled predictions vs proportion of individual deaths observed between 1995 and 2010 for each species. As such, $r^2$ measures the amount of species-level variation explained by the model or sub-models. By contrast, AUROC examines the model's ability to distinguish individuals that died from individuals that survived.")
@

We then examined the relative contribution species traits, random effects and carbon-dependent and independent hazards (Table.~\ref{table:1}) had on overall hazard rates by individually removing their effects and examining changes in both $r^2$ and AUROC. First, we found that the removal of census random effect $\delta_t$ had little impact on either $r^2$ or AUROC. The removal of either wood density or light demand resulted in similar magnitude effects on both $r^2$ and AUROC. By contrast, the removal of maximum dbh had a slightly smaller effect on both metrics. We detected a significant drop in both $r^2$ and AUROC when species random effects (i.e. the parameters that capture variation that is otherwise unexplained by traits) were removed, further highlighting that maximum dbh, light demand and wood density explain only a small fraction of the overall inter-specific variation observed among species. As expected, the removal of both species traits and species random effects resulted in a model that was unable to capture any inter-specific variation observed among species, and thus had low discriminatory ability. Lastly, we found the removal of the carbon-independent hazard term ($\gamma_s$) resulted in a lower $r^2$ relative to the removal of the carbon-dependent hazard ($\alpha_s e^{-\beta_s X_{i}}$), indicating that the former hazard explained more inter-specific variation. However, when examining the removal of these hazards in terms of the model's ability to correctly distinguish survivors from those that died (AUROC), we found the opposite -- that the removal of the carbon-dependent hazard resulted in substantially lower predictive ability.

\section*{Discussion}
Our Bayesian framework coupled with cross validation revealed that the most predictive hazard model of tropical tree death was one which: 1) partitioned mortality into carbon-dependent and carbon-independent hazards; 2) used stem diameter growth rather than basal-area growth; 3) included the effects of species' maximum dbh, wood density and light demand; and 4) incorporated temporal variability. Moreover, we found that rates of tropical tree mortality varied substantially between species and that the a species' maximum dbh, wood density and light demand explained only a fraction of the overall inter-specific variation.

The findings of this study provide empirical support for dynamic vegetation models that predict tree mortality as a function of both carbon-dependent and independent hazards \cite{Pacala:1996vm, Moorcroft:2001ws, McDowell:2011dr}. We show that regardless of growth measure (our proxy for carbon budget), incorporating both hazards significantly improves model predictive skill. This is because the carbon-dependent hazard allows for deaths associated with low carbon budgets, and as a consequence, incorporates intra-specific variability attributed to carbon related stresses (e.g. competition, parasites, herbivory). By contrast, the carbon-independent hazard accounts for deaths caused by events that arise irrespective of an individual's carbon budget (e.g. windthrow, lightning strike).

Additionally, the partitioning of mortality rates into both carbon-dependent and carbon-independent effects allowed us to not only estimate their relative contribution, but also how species traits affect each. Like many other studies \cite{King:1994vg, Ruger:2011cv, Kunstler:2016km}, our analyses highlighted the importance of light competition in influencing tropical tree demographic rates. Specifically, we found that that the removal of the carbon-dependent hazard term resulted in a substantial reduction in the model's ability to correctly distinguish survivors from those that died (Table~\ref{table:1}). This suggests deficiencies in carbon budgets are a major contributor to tree death on Barro Colorado Island.

Our analyses support previous research that advocated that wood density, species maximum size and light demand important explanatory variables of inter-specific variation in tropical rainforests \cite{Poorter:2006tq, Chave:2009iy}. Specificially, we show that light demanding species are not only more susceptible to both low growth rates and growth-independent hazards, but also exhibited shallower declines in growth-dependent hazards with increasing growth rates. By contrast, both high dbh or wood density species were found to be more tolerant of growth-independent hazards, corrobrating evidence reported elsewhere \cite{Chave:2009iy, Kraft:2010kq}. However, our analyses has also supports other work \cite{vanGelder:2006exa, Kunstler:2016km} which highlights that higher wood density species are also more tolerant of low growth rates, and thus, better able to tolerate shade and competiton.

While the effects of species' wood density, maximum size and light demand are now varyingly being incorporated into mortality algorithms of many vegetation models \cite{Moorcroft:2001ws,Scheiter:2013ed}, our analysis indicate that such effects are likely to only capture a small proportion of the overall inter-specific variation (Fig. \ref{fig4}, Table~\ref{table:1}). This means that such models are likely to severely underestimate the true variability in tropical rainforest mortality rates. As a consequence, this underestimation is likely to manifest in biased estimates of carbon, water and nutrient dynamics of ecosystems \cite{Johnson-2016}.

% Daniel I find these next three paragraphs as ho humm. I'm advocating for their deletion. Instead, we could just end with the above? Thoughts?

%Future research should focus on resolving how climate influences both carbon-dependent and carbon-independent hazards, particularly as climate-driven mortality is increasing \cite{vanMantgem:2009bi,Anderegg:2012ep,Mcdowell:2013wf}. Our analyses revealed a marginal decline in mortality at BCI from 1995 to 2010. Attributing temporal changes to particular climatic variables is challenging, however, due to low number temporal replication (n=3).

%We should also consider how the interval between censuses affects the estimation of carbon-dependent and carbon-independent hazards. Large census intervals may underestimate carbon-dependent mortality, as wide census intervals will not capture deaths due to rapid declines in growth, or events such as drought (although drought might also increase tree growth \cite{Condit2017}). Consequently, we may overestimate the relative contribution of carbon-independent hazards.

%Here we showcase a new framework for modelling tropical tree mortality that unifies empirical evidence from within and between species studies. This framework also provides an approach for partitioning mortality rates into carbon-dependent and carbon-independent hazards. Our findings reveal that while wood density is an important trait affecting mortality rates, we are still only capturing a fraction of the overall species variability in mortality rates.

\matmethods{
\vspace{-2em}
\subsection*{Data}
We derived plant mortality models using individual growth and survival data collected from a relatively undisturbed 50-ha tropical rainforest plot on BCI, Panama (9.15\textdegree;N, 79.85\textdegree;W). The climate on the island is warm and rainfall is seasonal with most falling between April and November \cite{Condit:2012nz}.

Within the 50-ha plot the diameter at breast height and survival status of all free-standing woody plants that were at least 1.3 m tall and had diameter $\geq$ 1 cm were recorded in 1981\textendash1983, 1985, and every 5 years thereafter \cite{Condit:2012nz}. For the purpose of modelling mortality as a function of past growth, we discarded data collected prior to 1990. This was because diameter measurements were rounded to the nearest 5 mm for individuals with dbh < 55 mm, whereas in later censuses all individuals were measured to the nearest millimetre \cite{Condit2017}. Consequently, we modelled tree mortality as a function of past growth (our measure of available carbon budget) for censuses 1995\textendash2000, 2000\textendash2005 and 2005\textendash2010. We discarded species that do not exhibit secondary growth (e.g. palms and ferns), contained fewer than 10 individuals or did not contain an estimate of wood density. We also excluded individuals that: 1) did not survive at least two censuses (two being required to estimate growth rate); 2) were not consistently measured at 1.3 m above ground; 3) were multi-stemmed; 4) resprouted or seemingly ``returned from the dead''; or 5) were extreme outliers -- stems which grew more than 5 cm yr$^{-1}$ or shrunk more than 25\% of their initial diameter. In total \Sexpr{n_obs(BCI_training_full)} observations were used in this study comprising \Sexpr{n_inds(BCI_training_full)} individual trees and \Sexpr{n_spp(BCI_training_full)} species. Because of computational costs, the models fit in this study do not include individual random effects, as this would require estimation of an additional \Sexpr{n_inds(BCI_training_full)} parameters. Instead, our models assume that repeat measurements of an individual are independent of one another. We believe this is a reasonable assumption given that there is approximately 5-years between censuses.

Wood density for each species was estimated by coring trees located within 15 km of the BCI plot \cite{Wright:2010fl}. Cores were broken into pieces, each 5 cm long and specific gravity of each piece was determined by oven drying (100\textdegree C) and dividing by the fresh volume (as measured by water displacement).

A species' light demand was calculated using annual canopy census data collected during 1985\textendash1990 and 1990\textendash1995. The canopy census recorded, in all 5 by 5 m subplots across the 50 ha plot, the presence of leaf in six height intervals (0-2,2-5,5-10,10-20,20-30, \textgreater 30 m). For each subplot, we calculated the number of strata > 2 m containing vegetation; and then transformed this to a light index ranging from 0 (dense shade) to 1 (gap). As light may penetrate into a subplot from the edge of a subplot, we rescaled this index to account for values in the eight immediate neighbouring subplots. Specifically, we used a weighted sum approach whereby the central subplot is assigned a weight of 8 and the eight neighbouring subplots are assigned a weight of 1. This meant that the contribution of the central plot was equivalent to the combined effect of all eight neighbouring plots. These weighted values were then summed and rescaled between 0 and 1 by dividing by the maximum value estimated across all subplots. The light demand for each species was estimated as the mean light index encountered by new saplings appearing in the census (Fig. S2).

A species' maximum dbh was determined from directly from the data, by taking the maximum dbh within the 95th percentile observed for each species.

\subsection*{Model fitting}

Eqs. \ref{eq:mortality_model}-\ref{eq:model} were fit to the data using Bayesian inference and with covariates for growth rate in previous census, species traits -- maximum dbh, light demand and wood density, as well as random effects. Growth rates were estimated from field measurements of diameter, which inevitably include observation error. In our dataset, \Sexpr{prop_neg_growth(BCI_training_full)}\% of estimated growth rates were negative. To ensure our mortality model was not biased by these unlikely values we first applied a probabilistic model to estimate \textit{``true growth''}, taking into account measurement error and the distribution of growth rate across the community (see Supplementary Material S1 for details; Fig. S1). The parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ were modelled as a function of species traits -- maximum dbh $\psi$, wood density, $\rho$, and light demand, $\upsilon$ -- and a species-level random effect:
\begin{equation}\label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \left(\frac{\rho}{0.6}\right)^{\alpha_{1}}\left(\frac{\upsilon}{0.7}\right)^{\alpha_{2}}\left(\frac{\psi}{15}\right)^{\alpha_{3}},
\end{equation}
with similar formulations for $\beta_s$ and $\gamma_s$. Here $\alpha_{1}$ is the effect of wood density $\rho$, $\alpha_{2}$ is the effect of light demand, $\upsilon$, and $\alpha_{3}$ is the effect of maximum  dbh, $\psi$. $\alpha_{0,s}$ captures any other species-level residual error that is otherwise not explained by these three species traits. These random effects were modelled as random realisations from log-normal distributions. The form of eq. \ref{eq:submodels} ensures that parameters remain positive; and on a log scale this equates to an additive linear model centered around the denominator. In this case we centered wood density on 0.6 g cm$^{-3}$, light demand on 0.7 and maximum dbh on 15 cm. We also centered growth rate $X_i$ at the lower 5\% quantile for both diameter increment and area growth (0.172 and 0.338, respectively), meaning $\alpha_s$ should be interpreted as the hazard rate when growth rate was very low. Weak priors on all hyper-parameters were set (see Supplementary Material S2 for details). Models were fit in R \Sexpr{paste(version$major, version$minor, sep='.')} using the package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{rstan2017} and employing some numerical optimisations (see Supplementary Material S3-S4 for details). We executed three independent chains and in all cases modelled parameters converged within 2000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \cite{Brooks:1998ju}. After discarding the first 1000 iterations as ‘burn in’, a further 1000 iterations were taken from the joint posterior. Species parameter estimates from the final model are shown in Figs. S3\textendash S5.

\subsection*{Evaluating model skill}
Predictive skill was quantified by estimating the average log loss across 10-folds for held-out data, $\bar{\mathcal{L}}$ (Fig. \ref{fig1}). Logarithmic loss -- commonly known as log loss, $\mathcal{L}$, measures the skill of a model by penalizing incorrect predictions, based on how wrong the predicted probability is from the observed outcome, $S_i$ (Fig. \ref{fig1}C). Lower $\mathcal{L}$ implies greater skill. The average log loss across all individuals for the $k$th fold of held-out data, $\mathcal{L}_k$, is then
\begin{equation} \label{eq:logloss}
\mathcal{L}_k = - \frac1{N_k}\,\sum_{i=1}^{N_k}\left(S_i\,\log(p_{i,t_2\rightarrow t_3}) + ( 1 - S_i)\,\log(1 - p_{i,t_2\rightarrow t_3})\right),
\end{equation}
where $N_k$ refers to the number of observations in fold.
}

\showmatmethods % Display the Materials and Methods section

\acknow{We thank H. Muller-Landau and anonymous reviewers for feedback and B. Carpenter for technical advice. We acknowledge S. Hubbell, R. Foster, R. Pérez, S. Aguilar, S. Lao, S. Dolins, \& hundreds of field workers for their contribution; and the National Science Foundation, Smithsonian Tropical Research Institute, \& MacArthur Foundation for funding the design, collection, quality control and management of long-term growth data at BCI. We also Fig. 1a uses images by Tracey Saxby, IAN Image Library. J.S.C., R.G.F., L.M. and D.S. were supported by the Science and Industry Endowment Fund (SIEF; RP04-174). D.S.F. and M.W. were supported by fellowships from the Australian Research Council.}

\showacknow % Display the acknowledgements section

\vspace{-5em}
\pnasbreak

% Bibliography
\bibliography{refs.bib}
%\pnasbreak
\end{document}